---
title: "In-class Exercise 9"
author: "Zhang Shujie (Suzy)"
date: "March 16, 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  warning: false
editor: visual
---

# Load packages

```{r}
pacman::p_load(tidygraph, ggraph, 
               visNetwork, lubridate, clock,
               tidyverse, graphlayouts)
```

# Importing network data from files

```{r}
GAStech_nodes <- read_csv("data/GAStech_email_node.csv")
GAStech_edges <- read_csv("data/GAStech_email_edge-v2.csv")
```

create 2 more columns

```{r}
GAStech_edges <- GAStech_edges %>%
  mutate(SendDate = dmy(SentDate)) %>%
  mutate(Weekday = wday(SentDate,
                        label = TRUE,
                        abbr = FALSE))
```

### **Wrangling attributes**

A close examination of *GAStech_edges* data.frame reveals that it consists of individual e-mail flow records. This is not very useful for visualisation.

In view of this, we will aggregate the individual by date, senders, receivers, main subject and day of the week.

The code chunk:

```{r}
GAStech_edges_aggregated <- GAStech_edges %>%
  filter(MainSubject == "Work related") %>%
  group_by(source, target, Weekday) %>%
    summarise(Weight = n()) %>%
  filter(source!=target) %>%
  filter(Weight > 1) %>%
  ungroup()
```

### **Using `tbl_graph()` to build tidygraph data model.**

In this section, you will use `tbl_graph()` of **tinygraph** package to build an tidygraph’s network graph data.frame.

Before typing the codes, you are recommended to review to reference guide of [`tbl_graph()`](https://tidygraph.data-imaginist.com/reference/tbl_graph.html)

```{r}
GAStech_graph <- tbl_graph(nodes = GAStech_nodes,
                           edges = GAStech_edges_aggregated, 
                           directed = TRUE)
```

```{r}
GAStech_graph %>%
  activate(edges) %>%
  arrange(desc(Weight))
```

### **Plotting a basic network graph**

The code chunk below uses [*ggraph()*](https://ggraph.data-imaginist.com/reference/ggraph.html), [*geom-edge_link()*](https://ggraph.data-imaginist.com/reference/geom_edge_link.html) and [*geom_node_point()*](https://ggraph.data-imaginist.com/reference/geom_node_point.html) to plot a network graph by using *GAStech_graph*. Before your get started, it is advisable to read their respective reference guide at least once.

```{r}
ggraph(GAStech_graph) +
  geom_edge_link() +
  geom_node_point()
```

### **Fruchterman and Reingold layout**

The code chunks below will be used to plot the network graph using Fruchterman and Reingold layout.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "kk") +
  geom_edge_link(aes()) +
  geom_node_point(aes())

g + theme_graph()
```

### **Modifying network nodes**

In this section, you will colour each node by referring to their respective departments.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes()) +
  geom_node_point(aes(colour = Department, 
                      size = 3))

g + theme_graph()
```

### **Modifying edges**

In the code chunk below, the thickness of the edges will be mapped with the *Weight* variable.

```{r}
g <- ggraph(GAStech_graph, 
            layout = "nicely") +
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 3)

g + theme_graph()
```

### **Working with *facet_edges()***

In the code chunk below, [*facet_edges()*](https://ggraph.data-imaginist.com/reference/facet_edges.html) is used. Before getting started, it is advisable for you to read it’s reference guide at least once.

```{r}
set_graph_style()

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)

g + facet_edges(~Weekday)
```

### **A framed facet graph**

The code chunk below adds frame to each graph.

```{r}
set_graph_style() 

g <- ggraph(GAStech_graph, 
            layout = "nicely") + 
  geom_edge_link(aes(width=Weight), 
                 alpha=0.2) +
  scale_edge_width(range = c(0.1, 5)) +
  geom_node_point(aes(colour = Department), 
                  size = 2)
  
g + facet_edges(~Weekday) +
  th_foreground(foreground = "grey80",  
                border = TRUE) +
  theme(legend.position = 'bottom')
```
